---
title: "week6_eda2020"
author: "Lucia M Rodriguez Bravo"
date: "March, 2020"
output:
  pdf_document:
    toc: yes
  html_document:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: console
---
# Graphing Data activity
## Data extraction
First we're calling the packages to be used
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
```

## Load data
Second, we're loading a database already included in R
```{r}
irisdata<-iris #I chose to create an object to look at the data more easily
# ------------------------------------------
#STEP 1: Did it load correctly?
head(irisdata) #loading correct
# ------------------------------------------
#STEP 2: Are the data types right?
sapply(irisdata,class) #classes correct
# ------------------------------------------
# STEP 3: Check for missing or impossible numeric values
range(irisdata$Sepal.Length) 
range(irisdata$Sepal.Width)
range(irisdata$Petal.Length)
range(irisdata$Petal.Width) #all good
# ------------------------------------------
# STEP 4: Check factor levels
levels(irisdata$Species)

# All OK
```

# Plot One Variable - Base
## Discrete variable
```{r}
#To plot counts (Y) of a discrete variable X (in this case species of flower) in Base
barplot(table(irisdata$Species)) #table creates a count
```

## Continuous variable
```{r}
#To plot a continuous variable X (in this case petal length) in Base
hist(irisdata$Petal.Length)

# Try choosing different breaks or bins of data
hist(irisdata$Petal.Length,breaks=4)

#A histogram with four columns
# Extract the hist function calculations
petalhist<-hist(irisdata$Petal.Length) #we create an object in which we can see our histogram function calculations 
petalhist
# A way to represent by species more clearly
boxplot(irisdata$Petal.Length~irisdata$Species)

```

# Plot One Variable - ggplot
## Discrete variable
```{r}
#To plot counts (Y) of a discrete variable X (in this case species of flower) in ggplot

Speciesgg<-ggplot(data = irisdata,aes(Species))+geom_bar()
Speciesgg #I chose to store the plot as an object

# Demonstrate how things can be layered (by creating a plot object and adding different types of plots
Speciesbase<-ggplot(data = irisdata,aes(Species)) 
Speciesbase #This is just the canvas
```

## Continuous variable
```{r}
#To plot a continuous variable X (in this case petal length) in ggplot
petalgg<-ggplot(data=irisdata,aes(Petal.Length))
petalgg
petalgg + geom_area(stat="bin") #warning just tells you that default 'bin=30'

# Follow the warning suggestion and explore different bin sizes
petalgg+geom_histogram(binwidth = 1) #sets width of the bins
petalgg+geom_histogram(bins = 3) #sets number of bins TO THREE. Vindication!

# Try to flip the axis order of your plot
petalgg+geom_histogram(binwidth = 1)+coord_flip() 
petalgg+geom_histogram(bins = 3)+coord_flip() 
# hashtagflipped B)
```

# Plot Two Variable - Base
## Continuous X, continuous Y
```{r}
# Explore different plots interchanging the 'predictive' and 'response' variables among the four available
#For predictive variable sepal width, response variable sepal length:
plot(Sepal.Length~Sepal.Width, data=irisdata)
#there are three species of flower in this dataset, 
#which messes with the ability to have a clear relation between variables. SO:
setosa<-filter(irisdata, Species=="setosa")
plot(Sepal.Length~Sepal.Width, data=setosa) #this one looks like there might be a relationship.
#Some other attempts:
plot(Petal.Length~Petal.Width, data=setosa)
plot(Petal.Width~Petal.Length, data=setosa)
#it looks like the width and length of the sepal has a more defined proportion than the width and length of the petal for this species

# Now I'll add a line representing linear model fit to the more successful graph:
plot(Sepal.Length~Sepal.Width, data=setosa)
abline(lm(Sepal.Length~Sepal.Width, data=setosa))
```

## Discrete X, continuous Y
```{r}
# Make a plot that most accurately represent the whole range of a continuous (Y) by species (X)

#IDONOTKNOW

```

# Plot Two Variable - ggplot
## Continuous X, continuous Y
```{r}

# Explore different plots interchanging the 'predictive' and 'response' variables among the four available-
#for this one I will plot petal length and sepal length
flowersize<-ggplot(data=irisdata, aes(x=Petal.Length, y=Sepal.Length))
flowersize+geom_point()
#now petal width and sepal width
flowerwidth<-ggplot(data=irisdata, aes(x=Petal.Width, y=Sepal.Width))
flowerwidth+geom_point() 
#not nearly as neat, it looks like different species might have different relationships
# Try adding lines representing different trend lines to help you explore patterns
# Add a linear model fit to your plots
#I'll do this to my more successful plot
flowersize+geom_point()+geom_smooth(method="lm", level=0.5)

# Add a 'Loess smoothing' fit to your plots
 flowersize+geom_point()+geom_smooth(span=100)
```

## Discrete X, continuous Y
```{r}
# Make a plot that most accurately represent the whole range of a continuous (Y) by species (X)
ggplot(data=irisdata,aes(x=Species,y=Petal.Length))+geom_boxplot()

# Try adding datapoints to this plot to reveal patterns not easy to see in a whiskerplot
ggplot(data=irisdata,aes(x=Species,y=Petal.Length))+geom_jitter()+geom_boxplot(alpha=0.7)
#Et voila
```





